---
title: "Project_2"
author: "Lee Panter, Arlin Tawzer, Nick Weaver"
date: "10/15/2018"
output: pdf_document
---
Markdown and Knitr options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Working directories
```{r}
P2WD="/Users/lee/Desktop/MATH_6388/Project_2"
setwd(P2WD)
```


Libraries & Packages:

- lubridate
```{r}
library(lubridate)
library(leaps)
library(car)
```

Data Dependencies:

- nonrookies.rda
  -- Note: this is a pre-altered dataset that has had rookies and several variables merged and removed.
```{r}
load(file = "/Users/lee/Desktop/MATH_6388/Project_2/nonrookies.rda")
dat=nonrookies
rm(nonrookies)
```

Initial Data Analysis
```{r}
(names=names(dat))
summary(dat)
```

Removal of non-quantitative predictors
```{r}
dat <- subset(dat, select = -c(yearID,teamID,lgID,playerID,
                                             birthYear,birthMonth,birthDay,
                                             birthCountry,birthState,birthCity,
                                             deathYear,deathMonth,deathDay,
                                             deathCountry,deathState,deathCity,
                                             nameFirst,nameLast,nameGiven,bats,
                                             throws,debut,finalGame,retroID,
                                             bbrefID,finalgame_year))
```

Set values of NA salary to 0, remove 0-salaried players
```{r}
for (i in 1:nrow(dat)){
  temp.sal.16 <- dat$salary_2016[i]
  temp.sal.15 <- dat$salary_2015[i]
  if (is.na(temp.sal.16))
  {dat$salary_2016[i] = 0}
  else if (is.na(temp.sal.15))
  {dat$salary_2015[i]=0}
}
dat=subset(dat, (salary_2016 != 0) & (salary_2015 != 0), select = index:GIDP)

rm(i); rm(temp.sal.15); rm(temp.sal.16)
```


Split data into training and test sets
```{r}
set.seed(123)
a <- 0.7*nrow(dat)
tmp.random<-sample(1:nrow(dat))
training<-dat[tmp.random[1:a],]
dim(training)

test=dat[-tmp.random[1:a],]
dim(test)
```

Create OLS model
```{r}
lmod.train.15=lm(salary_2015~.-salary_2016, data = training)
lmod.train.15.s=summary(lmod.train.15)
```

regsubsets
```{r}
regfit.train.15=regsubsets(salary_2015~.-salary_2016, data = training, nvmax = 21)
regfit.train.15.s=summary(regfit.train.15)
regfit.train.15.s$which
```

Define and Plot AIC
```{r}
n.train=dim(training)[1]
AIC=n.train*log(regfit.train.15.s$rss/n.train)+(1:20)*2
plot(AIC ~ I(1:20), ylab="AIC", xlab="Number of Predictors")

which.min(AIC) 
#The Number of predictors that minimizes AIC is 9
```


Define and Plot BIC
```{r}
BIC=regfit.train.15.s$bic
plot(BIC ~ I(1:20), ylab="BIC", xlab="Number of Predictors")
which.min(BIC)
#The number of predictors that minimizes BIC is 6
```


Define and Plot Cp
```{r}
cp=regfit.train.15.s$cp
plot(cp~I(1:20), ylab="Cp", xlab="Number of Predictors")
which.min(cp)
#The number of predictors that minimizes Cp is 9
```


Define New "Best" Functions according to the best 6 and 9 predictors

* Best six are: 

  - weight
  - G
  - AB
  - H
  - RBI
  - SH
  
* Best nine are:

  - weight
  - height
  - G
  - AB
  - H
  - RBI
  - HBP
  - SH
  - GIDP
  
```{r}
lm.Acp=lm(salary_2015~weight+G+AB+H+RBI+SH, data = training)
lm.BIC=lm(salary_2015~weight+height+G+AB+H+RBI+HBP+SH+GIDP, data=training)
```


Calculate Predicted values for test set
```{r}
pred.Acp.2015=predict.lm(lm.Acp, newdata = test)
pred.BIC.2015=predict.lm(lm.BIC, newdata = test)

SE.pred.Acp.2015=c()
SE.pred.BIC.2015=c()

for (i in 1:length(pred.Acp.2015))
{
  SE.pred.Acp.2015[i]=(test$salary_2015[i]-pred.Acp.2015[i])^2
  SE.pred.BIC.2015[i]=(test$salary_2015[i]-pred.BIC.2015[i])^2
}

(MSE.Acp.2015=sum(SE.pred.Acp.2015)/length(pred.Acp.2015))
(MSE.BIC.2015=sum(SE.pred.BIC.2015)/length(pred.BIC.2015))

#The smaller of the MSE measurements is for BIC (NOT that it is any good)

```


We now apply this model (BIC-selected variable model) towards prediction of the 2016 salaries, and calculate a loss function on this verification set:

```{r}
dat2=dat[,-3]
dat2$salary_2015=dat$salary_2016
pred.BIC.2016=predict.lm(lm.BIC, newdata = dat2)

SE.pred.BIC.2016=c()

for (i in 1:length(pred.BIC.2016))
{
  SE.pred.BIC.2016[i]=(dat$salary_2016[i]-pred.BIC.2016[i])^2
}

(MSE.BIC.2016=sum(SE.pred.BIC.2016)/length(pred.BIC.2016))

```



